---
title: "Data Cleaning -- Space Treasure"
author: "Paul A. Bloom"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(readr)
library(readxl)
#library(R.matlab)
```

This script cleans and preps data for further modeling -- both for 1-trial-back analyses and full RL models. 


# Load each subject's file in, then code stay/switch

```{r, echo=FALSE}
# read in file names from raw_data directory
files = list.files(path= '../raw_data/danl/', pattern = '.csv', full.names = TRUE)

# Load all files into 1 dataframe
space_data = files %>% 
    map_dfr(read_csv, col_types = cols(participantID= col_character()), .id = "source")

# Remove data for false start participant
space_data = dplyr::filter(space_data, !is.na(participantID))


# Recode participantID to be standardized
space_data = mutate(space_data,
                  participantID = parse_number(gsub('_V2W2', '', participantID))) %>%
  dplyr::filter(., !is.na(participantID))

# Check that each run had all 200 trials completed
participantN = space_data %>%
  filter(., choicePortal == 0 | choicePortal == 1) %>%
  dplyr::group_by(participantID) %>%
  dplyr::summarise(n  = n())
```

# Remove extra columns
```{r}
space_data = dplyr::select(space_data, -startTime, -date, -source)
```


# Code previous transition, previous reward, and stay-switch 
```{r}
# for early subjects before task code was edited -- fill in trialNum using the X column
space_data$trialNum = ifelse(is.na(space_data$trialNum), space_data$X, space_data$trialNum)

```

# Recode variables

```{r}
space_data = dplyr::select(space_data, 
                           id = participantID, trial = trialNum,
                           choice_1 = choicePortal,
                           choice_2 = choiceStage2,
                           reward, 
                           transition = transitionType,
                           rt_1 = rtStage1,
                           rt_2 = rtStage2, 
                           state = stage2Type,
                           everything())

space_data = mutate(space_data, choice_2 = ifelse(choice_2 %in% c(1,3), 0, 1),
                    state = state-1)
```


# Use `lag` function to get columns for previous decisions

```{r}
space_data = space_data %>%
  group_by(id) %>%
   mutate(., last_transition = lag(transition),
             last_reward = lag(reward),
             last_choice_1 = lag(choice_1),
             last_choice_2 = lag(choice_2),
             stay1 = ifelse(choice_1 == last_choice_1, 1, 0),
             stay2 = ifelse(choice_2 == last_choice_2, 1, 0)) %>%
  ungroup() %>%
  dplyr::mutate(state = case_when(
    choice_1 == 0 & transition == 'common' ~ 'A',
    choice_1 == 0 & transition == 'rare' ~ 'B',
    choice_1 == 1 & transition == 'common' ~ 'B',
    choice_1 == 1 & transition == 'rare' ~ 'A',
  ))
```

# Filter data for only healthy comparison participants

```{r}
demog = read_excel('../rawDemog/MASTER_W2_20200417.xlsx')
demog = dplyr::mutate(demog, participantID = parse_number(gsub('_V2', '', IDENT_SUBID))) %>%
  dplyr::select(id = participantID, age = SUBAGE_Session2, group = GROUP)

space_data = space_data %>%
  left_join(., demog) %>%
  dplyr::filter(group == 'C')


space_data_for_rl_model = space_data %>%
  dplyr::select(id, trial, choice_1, choice_2, reward, state) %>%
  mutate(choice_2 = ifelse(state=='B', choice_2+3, choice_2+1),
         choice_1 = choice_1 + 1)


danl_choice_1_matrix = space_data_for_rl_model %>%
  select(., id, choice_1, trial) %>%
  tidyr::spread(., key = id, value = choice_1) %>%
  select(., -trial) %>%
  t()

danl_choice_2_matrix = space_data_for_rl_model %>%
  select(., id, choice_2, trial) %>%
  tidyr::spread(., key = id, value = choice_2) %>%
  select(., -trial) %>%
  t()

danl_reward_matrix = space_data_for_rl_model %>%
  select(., id, reward, trial) %>%
  tidyr::spread(., key = id, value = reward) %>%
  select(., -trial) %>%
  t()


save(danl_choice_1_matrix, danl_choice_2_matrix, danl_reward_matrix, file = '../clean_data/danl_data_for_rl_model.rda')

```


# Write out cleaned data file
```{r}
write.csv(space_data, file = '../clean_data/danl_space_treasure_comps.csv', row.names = FALSE)
```



# Pull Hartley Lab Data Files


No stage2 data here
```{r}
# decker = read_csv('https://raw.githubusercontent.com/hartleylabnyu/online_two_step_replication/master/analysis_code_and_data/data/decker/decker_data.dat')
# decker_ages = read_csv('https://raw.githubusercontent.com/hartleylabnyu/online_two_step_replication/master/analysis_code_and_data/data/decker/decker_ages.csv') %>% 
#   dplyr::select(subj = subject_id, everything())
# decker = left_join(decker, decker_ages, by = c('subj'))
# 
# potter = read_csv('https://raw.githubusercontent.com/hartleylabnyu/online_two_step_replication/master/analysis_code_and_data/data/potter/potter_data.dat')
# potter_ages = read_csv('https://raw.githubusercontent.com/hartleylabnyu/online_two_step_replication/master/analysis_code_and_data/data/potter/potter_ages.csv') %>%
#   dplyr::select(subj = subject_id, everything())
# potter = left_join(potter, potter_ages, by = c('subj'))

```


```{r}
online = read_csv('https://raw.githubusercontent.com/hartleylabnyu/online_two_step_replication/master/analysis_code_and_data/data/online/online_mat_files/online_data_for_matlab.txt')

online_ages = read_csv('https://raw.githubusercontent.com/hartleylabnyu/online_two_step_replication/master/analysis_code_and_data/data/online/mbmf_ages.csv')

online = dplyr::left_join(online, online_ages, by = 'subject_id') %>%
  dplyr::select(trial_num = trial, everything(), -practice_trial) %>%
  dplyr::mutate(choice_1 = case_when(choice_1 == 2 ~ 1, choice_1 == 1 ~ 0, choice_1 == 0 ~ NA_real_),
                choice_2 = case_when(choice_2 == 2 ~ 1, choice_2 == 1 ~ 0, choice_2 == 0 ~ NA_real_),
                transition = case_when(
                  choice_1 == 0 & state == 2 ~ 'common',
                  choice_1 == 0 & state == 3 ~ 'rare',
                  choice_1 == 1 & state == 3 ~ 'common',
                  choice_1 == 1 & state == 2 ~ 'rare'),
                cohort = 'online')


# Decker data
decker_ages = read_csv('https://raw.githubusercontent.com/pab2163/online_two_step_replication/master/analysis_code_and_data/data/decker/decker_ages.csv')

decker = read_csv('https://raw.githubusercontent.com/pab2163/online_two_step_replication/master/analysis_code_and_data/data/decker/combined_decker_data.csv') %>%
  dplyr::mutate(choice_1 = case_when(choice1 == 2 ~ 1, choice1 == 1 ~ 0, choice1 == 0 ~ NA_real_),
                choice_2 = case_when(choice2 == 2 ~ 1, choice2 == 1 ~ 0, choice2 == 0 ~ NA_real_),
                transition = case_when(
                  choice_1 == 0 & state == 2 ~ 'common',
                  choice_1 == 0 & state == 3 ~ 'rare',
                  choice_1 == 1 & state == 3 ~ 'common',
                  choice_1 == 1 & state == 2 ~ 'rare'),
                cohort = 'decker',
                subject_id = parse_number(id)) %>%
  dplyr::select(subject_id, trial_num, choice_1, choice_2, transition, rt_1 = rts1, rt_2 = rts2, reward=money, cohort) 
  

decker = left_join(decker, decker_ages)


# Potter data

potter_ages = read_csv('https://raw.githubusercontent.com/pab2163/online_two_step_replication/master/analysis_code_and_data/data/potter/potter_ages.csv')

potter = read_csv('https://raw.githubusercontent.com/pab2163/online_two_step_replication/master/analysis_code_and_data/data/potter/combined_potter_data.csv') %>%
  dplyr::mutate(choice_1 = case_when(choice1 == 2 ~ 1, choice1 == 1 ~ 0, choice1 == 0 ~ NA_real_),
                choice_2 = case_when(choice2 == 2 ~ 1, choice2 == 1 ~ 0, choice2 == 0 ~ NA_real_),
                transition = case_when(
                  choice_1 == 0 & state == 2 ~ 'common',
                  choice_1 == 0 & state == 3 ~ 'rare',
                  choice_1 == 1 & state == 3 ~ 'common',
                  choice_1 == 1 & state == 2 ~ 'rare'),
                cohort = 'potter',
                subject_id = parse_number(id)) %>%
  dplyr::select(subject_id, trial_num, choice_1, choice_2, transition, rt_1 = rts1, rt_2 = rts2, reward=money, cohort) 
  

potter = left_join(potter, potter_ages)
```

Code lags in hartley data

```{r}
hartley = plyr::rbind.fill(online, decker, potter) %>%
  dplyr::select(id = subject_id, everything())


hartley = hartley %>%
    group_by(id) %>%
    mutate(., last_transition = lag(transition),
             last_reward = lag(reward),
             last_choice_1 = lag(choice_1),
             last_choice_2 = lag(choice_2),
             stay1 = ifelse(choice_1 == last_choice_1, 1, 0),
             stay2 = ifelse(choice_2 == last_choice_2, 1, 0)) %>%
    ungroup() 

# clean up trials not to be analyzed / generate matching choice numbers
hartley = hartley %>%
  dplyr::filter(!is.na(transition),  #no transition
                !is.na(last_transition),
                !is.na(rt_1),
                !is.na(rt_2),
                !is.na(choice_1),
                !is.na(choice_2)) %>% # first trial or trials after a skipped
                #!(cohort == 'online' & choice_2 ==0),
                #!(cohort == 'online' & choice_1 ==0)) %>% # no choice 2 %>%
  # dplyr::mutate(choice_2 = case_when(
  #   cohort == 'decker' ~ choice_2,
  #   cohort == 'online' & choice_2 == 1 ~ 0,
  #   cohort == 'online' & choice_2 == 2 ~ 1,
  # ),
  # choice_1 = case_when(
  #   cohort == 'decker' ~ choice_1,
  #   cohort == 'online' & choice_1 == 1 ~ 0,
  #   cohort == 'online' & choice_1 == 2 ~ 1),
  dplyr::mutate(state = case_when(
    choice_1 == 0 & transition == 'common' ~ 'A',
    choice_1 == 0 & transition == 'rare' ~ 'B',
    choice_1 == 1 & transition == 'common' ~ 'B',
    choice_1 == 1 & transition == 'rare' ~ 'A',
  ),
  group = case_when(age < 13 ~ 'child',
                            age >= 13 & age <= 17 ~ 'adol',
                            age > 17 ~ 'adult'),
  rt_1 = ifelse(cohort == 'online', rt_1, rt_1*1000),
  rt_2 = ifelse(cohort == 'online', rt_2, rt_2*1000),
  )




hartley %>%
  group_by(id, cohort, last_transition, last_reward, group) %>%
  summarise(n = n(), 
            stay1_prop = sum(stay1)/n) %>%
  ggplot(aes(x = last_transition, y = stay1_prop, color = factor(last_reward))) +
  stat_summary(fun.data = mean_cl_boot) +
  facet_grid(rows = vars(group), cols = vars(cohort, last_reward))

space_data %>%
  group_by(id, last_transition, last_reward) %>%
  summarise(n = n(), 
            stay1_prop = sum(stay1)/n) %>%
  ggplot(aes(x = last_transition, y = stay1_prop)) +
  stat_summary(fun.data = mean_cl_boot) +
  facet_grid(~last_reward)

write.csv(hartley, file = '../clean_data/hartley_lag.csv', row.names = FALSE)

```
